Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-02-23 ThesisJP/Background/background.tex
Writing background...

In this chapter the background information needed in order to understand the topics of this thesis and its papers is given. First, in Section \ref{sec:gpgpu}, general-purpose programming on graphics processing units (\nom{GPGPU}{General-purpose programming on graphics processing units}) is reviewed, followed by an introduction to medical ultrasound imaging in Section \ref{sec:ultrasound}. Finally, the concepts of volume rendering and ultrasound field simulations are presented in Section \ref{sec:volren} and \ref{sec:field}.

\section{General-purpose computing on graphics processing units}\label{sec:gpgpu}
In a recent essay, Herb Sutter, a leading authority on software development, reviews the current state of the software and computer industry \cite{HerbSutter}. The essay bears the name "Welcome to the Jungle" and is a sequel to the essay "The Free Lunch is Over" from 2004 \cite{HerbSuttera}. The two essays spins around the fact that CPU manufactures hit a frequency-wall in the beginning of the 21st century. Until then, all mainstream computer programs where typically running in a single thread on a single CPU \textit{processing core}, and a programmer could expect the software to annually gain performance without touching the code. This was the "free-lunch"-era as depicted in Fig.\,\ref{fig:jungle}. The increase in processing power came mostly from an ever-growing clock frequency. However, power consumption and heat generation where also growing at the same rate, until the level of cooling required finally became to much in 2004.  From that point, chip manufactures had to pursue a different direction  than just increasing the clock frequency. The immediate solution was to embed several cores in one CPU chip. This insured continuation of Moore's law, since the number of transistors per chip could continue to grow. For programmers, multi-core CPUs meant that computer programs now had to be multi-threaded in order to annually gain performance. It was a concurrency revolution, as Herb Sutter named it, and  the "free lunch" was over.

@@

Around the same time as CPU manufactures hit the frequency wall, people started experimenting with programmable shaders which recently had been introduced in the field of graphics programming. The graphics processing pipeline had before this been a fixed function pipeline, where e.g. geometry and textures where fed to the graphics processing unit (GPU) in one end and a rasterized rendered image was outputted in the other. What programmable shaders  introduced was the ability to replace some steps in the fixed-function pipeline with custom computations. This made it possible to utilize the GPU for other tasks than just rendering graphics \cite{Seland2007}. The rationale behind this exploration was that GPUs were already highly available in consumer PCs, and hidden behind its graphics interface were processing powers an order of magnitude larger than what the CPU could provide (Fig.\,\ref{fig:cpu_vs_gpu}). Actually, at the time when CPU manufacturers hit the frequency wall the GPU had already gone multi-core, first of all driven by an ever-increasing demand for more realistic computer games. Second, the rendering pipeline exhibits minimal execution path branching, and data are typically used once. Designers of GPUs could therefore skip advance feature, found in most CPUs, like branch prediction and different levels of data caching. These measures reduced the silicon footprint of each core and made it possible to add multiple cores to one GPU before CPU designers where able to do the same. Today, even if more advanced caching has been added to GPUs, this is still the main reason why GPUs have higher peak performance than CPUs (Fig.\,\ref{fig:cpu_vs_gpu}). Another reason is the inherent parallel nature of the rendering problem. When geometry is projected and rasterized the same instruction is typically needed for a lot of data at the same time. This makes a special kind of instruction, Single Instruction Multiple Data (\nom{SIMD}{Single instruction multiple data (instruction)}) instruction \cite{Flynn1966}, especially suited for this kind of problem. The differences between how SIMD instructions are utilized in modern CPUs and GPUs are discussed in Section \ref{sec:cpu_vs_gpu}.

For the first adopters of GPU programming, or General-purpose programming on graphics processing units (GPGPU) as it was named, it soon became evident that offloading all types of computations to the GPU was not always a good idea. In most cases it is important to balance the load equally between the CPU and GPU to obtain their combined computationally power. Balancing typically means that highly parallel and computationally intense tasks should be offloaded to the GPU and serialized and memory intensive work should stay on the CPU. This paradigm of utilizing specialized cores for solving specific problems is today known as heterogeneous computing. The specialized core can be a GPU or e.g. a field-programmable gate array (\nom{FPGA}{Field-programmable gate array}). A recent trend is that specialized accelerator cores are embedded onto the CPU. This heterogeneous processors is often referred to as an accelerated processing unit (\nom{APU}{Accelerated processing unit}). The last three generations of CPUs from Intel have all had on-chip integrated graphics. Even if an on-chip GPU is less powerful than a high-performance s discrete GPU, valuable time is saved by not having to send data across the PCI Express bus, which connects the different parts of a computer to the motherboard. This is the reason why one of the first rules of GPGPU (with discrete GPUs) is to minimize CPU-to-GPU memory transfers.

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{img/CPU-GPU.pdf}
\caption{Schematic overview of CPU (Intel Haswell) and GPU (Nvidia Kepler) architecture. Note how more space is dedicated for compute cores on a GPU, and how more space is dedicated to control and caching on a CPU.}
\label{fig:cpu_gpu}
\end{figure}

In Section \ref{sec:gpgpu} we introduced a special type of instructions, namely SIMD instructions, which are designed to execute a single instruction across multiple data element in one clock cycle. At the end of the 20th century, Intel introduced a SIMD instruction set for their Pentium III processor called the Streaming SIMD Extensions (\nom{SSE}{Streaming SIMD extensions (instruction)}). This extension has seen several updates since. The first SEE instructions set had 128 bit long registers which meant that four 32 bit floating point values could be processed per clock cycle\footnote{For simplicity we will assume that the whole register is processed in one clock cycle even if the actual implementation might distributes one instruction for the register across multiple cycles.}. Updates of the SSE instruction set have added new instructions (SSE2, SSE3, and SSE4) and longer registers (\nom{AVX}{Advanced vector extensions (instruction)}, AVX2, and AVX-512). AVX and AVX2 adds 256 bit long registers and fused multiply-accumulate instructions (\nom{FMA}{Fused multiply-accumulate (instruction)})  which means that sixteen 32 bit floating point values can be processed per clock cycle per core. A new extension, AVX-512, extends the registers to 512 bit and is found in the Intel accelerator board  Xeon Phi and might show up in the next generation of CPUs. The latest CPU architecture from Intel, the Haswell architecture, has two processing ports supporting FMA-AVX2 instructions, and is therefore capable of processing 32, 32 bit, floating point values per clock cycle per core (Fig.\,\ref{fig:cpu_gpu}). For an eight core\footnote{An 8-core Haswell CPU is scheduled to launch during 2014.} CPU with a clock frequency of 3.8 GHz this sums to 
\begin{equation}
32*8*3.8 = 972.8\,\text{GFLOPS},
\end{equation}
hence a theoretical peak performance of close to one tera single precision floating point operations per second (TFLOPS). This is the same number as visualized in Fig\,\ref{fig:cpu_vs_gpu} for the Haswell architecture. In addition, the most powerful integrated GPU found on Haswell CPUs have 40 execution units capable of processing sixteen, 32 bit, floating points at 1.1 GHz. This adds
\begin{equation}
16*40*1.1 = 704\,\text{GFLOPS}
\end{equation}
in extra processing power to the CPU chip.

Modern GPUs can also be interpreted as a unit consisting of multiple SIMD processors (Fig.\,\ref{fig:cpu_gpu}). The latest GPU architecture by Nvidia have up to 15 SIMD units capable of processing 192, 32 bit, floats per clock cycle at 745 MHz. These GPUs are also capable of performing FMA instructions, effectively doubling the number of theoretical FLOPS. The theoretical peak single precision FLOPS for this architecture is by that found to be 
\begin{equation}
15*192*2*745 = 4291\,\text{GFLOPS}.
\end{equation} 
Each SIMD core also comprise 16 special function units (\nom{SFU}{Special function unit (sin, exp, sqrt, ect.)}) which add close to 1 TFLOPS in additional computational power. This adds up to five TFLOPS in computationally power as shown in Fig.\,\ref{fig:cpu_vs_gpu} for the GeForce GTX 780 TI GPU. From Fig.\,\ref{fig:cpu_gpu} we see how the high number of SIMD cores found on GPUs and their larger width results in five times the computational performance of a high-end CPU (without the integrated GPU).

Note that the derived numbers are strictly theoretical and that actual throughput is highly algorithm dependent. Nevertheless, in theory, the benefit of GPU computing over CPU computing is currently a factor of five and ten\footnote{Until an eight core Haswell CPU is launched later this year.} in raw computationally power. A claimed speed-up larger than this is typically caused by comparing a non-optimized CPU implementation with an optimized GPU implementation or by comparing uneven hardware \cite{Lee2010, Kothapalli2013}. Also note that in this section we have focused on CPUs from Intel and GPUs from Nvidia, however, CPUs and GPUs from AMD have similar specifications.

@@

\subsection{Programming a GPU}
Early GPGPU programming required expert knowledge about computer graphics and the remaining fixed functions in the pipeline. When GPGPU gained interest it quickly became evident that new programing languages where needed.  One of the first GPGPU oriented languages for the GPU was the Brook language by Buck \textit{et al.} \cite{Buck2004}. Buck later joined Nvidia where he led the work on \nom{CUDA}{Compute unified device architecture (GPGPU language)}, a GPGPU language and framework by Nvidia for Nvidia GPUs only. The first version of CUDA was released in 2007. The following year, Apple launched a GPGPU framework called \nom{OpenCL}{Open Compute Language (GPGPU language)}, a standard which are maintained by the Khronos Group. These two are still today the main programming frameworks for GPGPU. CUDA only runs on Nvidia GPUs while OpenCL can now run on GPUs from both Nvidia and AMD, CPUs from both Intel and AMD, and on FPGAs from Altera.

@@

Both CUDA and OpenCL is centered around a kernel function which are executed be a large group of threads. The kernel function, for both CUDA and OpenCL, has a C-like syntax with some additional specifiers. A CUDA kernel is depicted in Fig.\,\ref{fig:gpu_kernel}. The \texttt{\_\_global\_\_} specifier tells the CUDA compiler that this function is a kernel function, and, in CUDA, this kernel function is launched by using a CUDA specific syntax:
\begin{align*}
\texttt{square<<<grid, block>>>(a, N)}
\end{align*}
where \texttt{grid} and \texttt{block} is a structure containing the grid and block size as depicted in Fig.\,\ref{fig:gpu_grid}, and \texttt{a} is a pointer to a memory buffer located on the GPU. When launched, a copy of the kernel is executed for each thread in each block. Each kernel get as input its location in the compute grid (\texttt{blockIdx} and \texttt{threadIdx}).

C++ AMP.

Architecture and how to program. See Paper II.

It it is also much less costly to switch between a pixel-thread for the GPU then a program thread for the CPU.

Hopefully, in the future, GPU programming will be the job of e.g. c++ compiler, and direct fiddling with GPU kernels should only be needed if maximum throughput is absolutely needed. (In the same way the compiler is able to translate c++ code into SSE and AVX instructions.)

\subsection{Accelerators in medical ultrasound imaging }
Add sentence about siemens and supersonic imagine scanners. They use GPUs? What does this mean? FPGAs and ASICs.

@@

\subsection{Cardiac anatomy}
Small section about heart anatomy.

@@

\subsection{Basic beamforming}

>>>>

In this chapter the background information needed in order to understand the topics of this thesis and its papers is given. First, in Section \ref{sec:gpgpu}, general-purpose computing on graphics processing units (\nom{GPGPU}{General-purpose computing on graphics processing units}) is reviewed, followed by an introduction to medical ultrasound imaging in Section \ref{sec:ultrasound}. Finally, the concepts of volume rendering and ultrasound field simulations are presented in Section \ref{sec:volren} and \ref{sec:field}.

\section{General-purpose computing on graphics processing units}\label{sec:gpgpu}
In a recent essay, Herb Sutter, a leading authority on software development, reviews the current state of the software and computer industry \cite{HerbSutter}. The essay bears the name "Welcome to the Jungle" and is a sequel to the essay "The Free Lunch is Over" from 2004 \cite{HerbSuttera}. The two essays spins around the fact that manufactures of central processing units (CPU) hit a frequency-wall in the beginning of the 21st century. Until then, all mainstream computer programs where typically running in a single thread on a single CPU \textit{processing core}, and a programmer could expect the software to annually gain performance without touching the code. This was the "free-lunch"-era as depicted in Fig.\,\ref{fig:jungle}. The increase in processing power came mostly from an ever-growing clock frequency. However, power consumption and heat generation where also growing at the same rate, until the level of cooling required finally became to much in 2004.  From that point, chip manufactures had to pursue a different direction  than just increasing the clock frequency. The immediate solution was to embed several cores in one CPU. This insured continuation of Moore's law, since the number of transistors per chip could continue to grow. For programmers, multi-core CPUs meant that computer programs now had to be multi-threaded in order to annually gain performance. It was a concurrency revolution, as Herb Sutter named it, and  the "free lunch" was over.

@@

Around the same time as CPU manufactures hit the frequency wall, people started experimenting with programmable shaders which recently had been introduced in the field of graphics programming. The graphics processing pipeline had before this been a fixed function pipeline, where e.g. geometry and textures where fed to the graphics processing unit (GPU) in one end and a rasterized rendered image was outputted in the other. What programmable shaders  introduced was the ability to replace some steps in the fixed-function pipeline with custom computations. This made it possible to utilize the GPU for other tasks than just rendering graphics \cite{Seland2007}. The rationale behind this exploration was that GPUs were already highly available in consumer PCs, and hidden behind its graphics interface were processing powers an order of magnitude larger than what the CPU could provide (Fig.\,\ref{fig:cpu_vs_gpu}). Actually, at the time when CPU manufacturers hit the frequency wall the GPU had already gone multi-core, first of all driven by an ever-increasing demand for more realistic computer games. Second, the rendering pipeline exhibits minimal execution path branching, and data are typically used once. Designers of GPUs could therefore skip advance feature, found in most CPUs, like branch prediction and different levels of data caching. These measures reduced the silicon footprint of each core and made it possible to add multiple cores to one GPU before CPU designers where able to do the same. Today, even if more advanced caching has been added to GPUs, this is still the main reason why GPUs have higher peak performance than CPUs (Fig.\,\ref{fig:cpu_vs_gpu}). Another reason is the inherent parallel nature of the rendering problem. When geometry is projected and rasterized the same instruction is typically needed for a lot of data at the same time. This makes a special kind of instruction, the Single Instruction Multiple Data (\nom{SIMD}{Single instruction multiple data (instruction)}) instruction \cite{Flynn1966}, especially suited for this kind of problem. The differences between how SIMD instructions are utilized in modern CPUs and GPUs are discussed in Section \ref{sec:cpu_vs_gpu}.

For the first adopters general-purpose computing on GPUs (GPGPU) it soon became evident that offloading all types of computations to the GPU was not always a good idea. In most cases it is important to balance the load equally between the CPU and GPU to obtain their combined computationally power. Balancing typically means that highly parallel and computationally intense tasks should be offloaded to the GPU and serialized and memory intensive work should stay on the CPU. This paradigm of utilizing specialized cores for solving specific problems is today known as heterogeneous computing. The specialized core can be a GPU or e.g. a field-programmable gate array (\nom{FPGA}{Field-programmable gate array}). A recent trend is that specialized accelerator cores are embedded onto the CPU. This heterogeneous processors is often referred to as an accelerated processing unit (\nom{APU}{Accelerated processing unit}). The last three generations of CPUs from Intel have all had on-chip integrated graphics. Even if an on-chip GPU is less powerful than a high-performance discrete GPU, valuable time is saved by not having to send data across the PCI Express bus, which connects the different parts of a computer to the motherboard. This is the reason why one of the first rules of GPGPU (with discrete GPUs) is to minimize CPU-to-GPU memory transfers.

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{img/CPU-GPU.pdf}
\caption{Schematic overview of CPU (Intel Haswell) and GPU (Nvidia Kepler) architecture. Note how more space is dedicated to compute cores on a GPU, and how more space is dedicated to control and caching on a CPU.}
\label{fig:cpu_gpu}
\end{figure}

In Section \ref{sec:gpgpu} we introduced a special type of instructions, namely SIMD instructions, which are designed to execute a single instruction across multiple data element in one clock cycle. At the end of the 20th century, Intel introduced a SIMD instruction set for their Pentium III processor called the Streaming SIMD Extensions (\nom{SSE}{Streaming SIMD extensions (instruction)}). This extension has seen several updates since. The first SEE instructions set had 128 bit long registers which meant that four, 32 bit, floating point values could be processed per clock cycle\footnote{For simplicity we will assume that the whole register is processed in one clock cycle even if the actual implementation might distributes one instruction for the register across multiple cycles.}. Updates of the SSE instruction set have added new instructions (SSE2, SSE3, and SSE4) and longer registers (\nom{AVX}{Advanced vector extensions (instruction)}, AVX2, and AVX-512). AVX and AVX2 adds 256 bit long registers and fused multiply-accumulate instructions (\nom{FMA}{Fused multiply-accumulate (instruction)})  which means that sixteen 32 bit floating point values can be processed per clock cycle per core. A new extension, AVX-512, extends the registers to 512 bit and is found in the Intel accelerator board  Xeon Phi and might show up in the next generation of CPUs. The latest CPU architecture from Intel, the Haswell architecture, has two processing ports supporting FMA-AVX2 instructions, and is therefore capable of processing 32, 32 bit, floating point values per clock cycle per core (Fig.\,\ref{fig:cpu_gpu}). For an eight core\footnote{An 8-core Haswell CPU is scheduled to launch during 2014.} CPU with a clock frequency of 3.8 GHz this sums to 
\begin{equation}
32*8*3.8 = 972.8\,\text{GFLOPS},
\end{equation}
hence a theoretical peak performance of close to one tera single precision floating point operations per second (TFLOPS). This is the same number as visualized in Fig\,\ref{fig:cpu_vs_gpu} for the Haswell architecture. In addition, the most powerful integrated GPU found on Haswell CPUs have 40 execution units capable of processing sixteen, 32 bit, floating points at 1.1 GHz. This adds
\begin{equation}
16*40*1.1 = 704\,\text{GFLOPS}
\end{equation}
in extra processing power to the APU.

Modern GPUs can also be interpreted as a unit consisting of multiple SIMD processors (Fig.\,\ref{fig:cpu_gpu}). The Kepler architecture by Nvidia have up to 15 SIMD units capable of processing 192, 32 bit, floats per clock cycle at 745 MHz. These GPUs are also capable of performing FMA instructions, effectively doubling the number of theoretical FLOPS. The theoretical peak single precision FLOPS for this architecture is by that found to be 
\begin{equation}
15*192*2*745 = 4291\,\text{GFLOPS}.
\end{equation} 
Each SIMD core also comprise 16 special function units (\nom{SFU}{Special function unit (sin, exp, sqrt, ect.)}) which add close to 1 TFLOPS in additional computational power. This adds up to a total of five TFLOPS in computationally power as shown in Fig.\,\ref{fig:cpu_vs_gpu} for the GeForce GTX 780 TI GPU. From Fig.\,\ref{fig:cpu_gpu} we see how the high number of SIMD cores found on GPUs and their larger width results in five times the computational performance of a high-end CPU (without the integrated GPU).

Note that the derived numbers are strictly theoretical and that actual throughput is highly algorithm dependent. Nevertheless, in theory, the benefit of GPU computing over CPU computing is currently a factor of five or ten\footnote{Until an eight core Haswell CPU is launched later this year.} in raw computationally power. A claimed speed-up larger than this is typically caused by comparing a non-optimized CPU implementation with an optimized GPU implementation or by comparing uneven hardware \cite{Lee2010, Kothapalli2013}. Also note that in this section we have focused on CPUs from Intel and GPUs from Nvidia, however, CPUs and GPUs from AMD have similar specifications.

@@

\subsection{Programming a GPU}
Early GPGPU programming required expert knowledge about computer graphics and the remaining fixed functions in the pipeline. When GPGPU gained interest it quickly became evident that new programing languages where needed.  One of the first GPGPU oriented languages for the GPU was the Brook language by Buck \textit{et al.} \cite{Buck2004}. Buck later joined Nvidia where he led the work on \nom{CUDA}{Compute unified device architecture (GPGPU language)}, a GPGPU language and framework by Nvidia for Nvidia GPUs only. The first version of CUDA was released in 2007. The following year, Apple launched a GPGPU framework called \nom{OpenCL}{Open Compute Language (GPGPU language)}, a standard which are maintained by the Khronos Group. These two frameworks are still today the main programming frameworks for GPGPU. Where CUDA only runs on Nvidia GPUs, OpenCL can now run on GPUs from both Nvidia and AMD, CPUs from both Intel and AMD, FPGAs from Altera, and mobile processors from \nom{ARM}{Designer of mobile chip architectures}.

@@

Both CUDA and OpenCL is centered around a kernel function which are executed by a large group of threads. The kernel function, for both CUDA and OpenCL, has a C-like syntax with some additional specifiers. A CUDA kernel is depicted in Fig.\,\ref{fig:gpu_kernel}. The \texttt{\_\_global\_\_} specifier tells the CUDA compiler that this function is a kernel function, and in CUDA, this kernel function is launched by using a CUDA specific syntax:
\begin{align*}
\texttt{square<<<grid, block>>>(a, N)},
\end{align*}
where \texttt{grid} and \texttt{block} is 3-component vectors specifying grid and block size in "xyz" as depicted in Fig.\,\ref{fig:gpu_grid}, \texttt{a} is a pointer to a memory buffer located on the GPU, and \texttt{N}$*$\texttt{N} is the buffer size. When launched, a copy of the kernel is scheduled  for each thread in each block, and each kernel get as input its location in the compute grid (\texttt{blockIdx} and \texttt{threadIdx}) together with the grid size (\texttt{gridDim} and \texttt{blockDim}). This information is then typically used to calculate which memory elements to gather from the global memory buffer \texttt{a}. In order to formulate a given problem as a kernel function the global memory input to each thread must not rely on the output from other threads, hence it must be possible for the GPU to scheduled threads in any order.

This is of course a minimal example of GPU programing. A full example will involve several lines of code transferring data to and from the GPU, and the kernel could include manual administration of the L1 cache (shared memory), and intra-block thread barriers. This is some of the reasons why writing a high performance GPU kernel is too often found to be a complex and time consuming task.

No doubt, it is still a major overhead involved when writing GPU code, and in the future we will probably see work aimed at reducing this overhead.  A recent development is the introduction of C++ AMP which enables developers to write accelerated code (e.g. by a GPU) in plain C++. Hopefully, in the future, GPU programming will be a job for compilers, and direct fiddling with GPU kernels should only be needed if maximum throughput is absolutely needed. This is similar to how SIMD instructions is handled in CPU code today. Most of the time developers will rely on the CPU compilers to analyze the code and auto-apply SIMD instruction whenever it is possible. However, when maximum throughput is needed, developers with deep knowledge of the problem at hand are often able to make faster code by manually writing SIMD instruction code. The relevance of auto-applied GPU instructions by compilers will also increasing due to the recent embedding of GPUs on the same chip as the CPU. Soon, when writing code, we will probably talk about coding a heterogeneous core where some operations are automatically offloaded to special cores (e.g. a GPU) by the compiler.

@@

@@

Small section about heart anatomy.

\subsection{Accelerators in medical ultrasound imaging}
Ultrasound imaging systems generates an incredible amount of data. A modern system capable of 3D imaging will generate several tera bytes of data in the front end of the system during a clinical scan. The data rate from a 2D scan is "only" in order giga bytes per second. To handle these amount of data, ultrasound imaging systems have relied heavily on application-specific integrated circuits (\nom{ASIC}{Application-specific integrated circuit}) for all steps in the processing pipeline. However, as the computationally power of CPUs and GPUs have increased, ASICs, e.g. for scan conversion, have gradually been converted into software \cite{Guracar2013}. One of  the last remaining ASICs is the beamformer, whose job is to steer and focus the ultrasound beam in space. This operation involves a reduction of the data rate with a factor equal to the number of elements in the ultrasound probe. The input data rate is therefor very high. Nevertheless, real-time beamforming has recently been achieved on a GPU \cite{Song2012}, and a French company (SuperSonic Imagine) are utilizing GPUs in their plane wave ultrasound beamforming algorithm \cite{Tanter2014}. Beamforming is therefore finally moving into software to. The largest players in the medical ultrasound business, General Electric and Philips, are also well aware of where the trend is going \cite{Thomenius2012} \cite{Metz2011}. Software beamforming opens up new possibilities for advanced processing. One example is Capon beamforming which is explored in this thesis (\textbf{Paper\,I} and \textbf{II}), another is advanced 3D visualization as presented in \cite{solteszova2010multidirectional}, \textbf{Paper\,IV}, and \textbf{IX}.
